Prediction Developing Countries' Life Expectancy

### 2025-02-20 paper.qmd
did preliminary results code/graphs

---
title: "Prediction Developing Countries' Life Expectancy"
author: 
  - Siddharth Gowda
  - Syed Hasan
  - Arshh Relan
date: today
date-format: long
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

ggplot(data_interest, aes(x = alcohol, fill = alcohol)) +
  geom_bar() +
  theme_minimal() +
  labs(title = "Number of Countries that Consume Alcohol", 
       x = "Alcohol Category", y = "Number of Countries") +
  scale_fill_manual(values = c("Negligible" = "skyblue", "Consumed" = "salmon"))

@@

>>>>

---
title: "Prediction Developing Countries' Life Expectancy"
author: 
  - Siddharth Gowda
  - Syed Hasan
  - Arshh Relan
date: today
date-format: long
format: pdf
number-sections: true
toc: true
---

@@

ggplot(data_interest, aes(x = alcohol, fill = alcohol)) +
  geom_bar(color = "black") +
  theme_minimal() +
  labs(title = "Number of Countries that Consume Alcohol", 
       x = "Alcohol Category", y = "Number of Countries") +
  scale_fill_manual(values = c("Negligible" = "skyblue", "Consumed" = "salmon"))

@@

## Basic Linear Model Summary

Model is Life Expectancy = GDP Per Capita + Infant Deaths (per 1000 people) + HIV/AID Deaths (per 1000 live births) + Health Expenditure (as a percentage of GDP per Capita) + Average Number of Years of Schooling + Alcohol Consumption (Negligible Consumption vs. Alcohol is Consumed)

```{r}
#| warning: false
#| message: false
#| echo: false

preliminary_model = lm(life_expectancy ~ infant_deaths + hiv_aids + gdp_capita + 
                         schooling + percentage_expenditure + alcohol, 
                       data=data_interest)

#summary(preliminary_model)
modelsummary(preliminary_model)

## Residual Plots

```{r}
#| label: fig-resplots
#| layout-ncol: 3
#| fig-cap: ""
#| fig-subcap: ["", "", ""]
#| fig-align: "center"
#| echo: false
#| warning: false
#| message: false
#| fig.show: "hold"

plot(preliminary_model, which = 1)
plot(preliminary_model, which = 2)
standard_residuals = rstandard(preliminary_model)
plot(data_interest$gdp_capita, standard_residuals, main="Standardized Residuals V. GDP Per Capita", 
     xlab="GDP Per Capita", ylab="Standardized Residuals")

### Linear Regression Residual Assumptions:

#### Linearity
#### Normal Errors
#### Constant Variance Assumption

# Conclusion
